---
title: " "
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
  rmdformats::material:
    self_contained: true

---

<br><br><br><br><br>
```{r include=FALSE}
library(rmdformats)
```
***
## Soutions
***
<br><br>
***
## The Basics
***

1) You can use any variable name to show that 23 is only divisible by itself and 1 but 22 is divisible by other numbers, eg:

```{r}
Not_Prime <- 22
Prime <- 23

Not_Prime / 2
Prime / 2

```

Of course, in R there are a thousand ways to skin a cat. You can also use the numeric operator %%, which checks the remainder left after a division:

```{r}
Not_Prime %% 2
Prime %% 2

```

We'll talk about *functions* later on, but of course with R if you want something done, there is probably an appropriate function simply one google away. I googled and found the funtion primes, which directly checks if something is a primer number. After installing, I simply used the is_prime function:

```{r}
library(primes)
is_prime(Prime)
is_prime(Not_Prime)
```

2) You can use an expression from the last exercise! Think about types of data - numerical, character, logical, integer:

```{r}
this_is_logical <- is_prime(Not_Prime)
this_is_numerical <- 4
this_is_character <- "hi there!"
this_is_integer <- 1:24

class(this_is_logical)
class(this_is_numerical)
class(this_is_character)
class(this_is_integer)
```

Of course, there *are* more datatypes. A google might tell you about complex data, which may remind you of high schol algebra - it's where you can use imaginary numbers to replace data:

```{r}
this_is_complex = 2i + 1
class(complex)

```

